### 2021-05-23 assertions/doc/thesis-bernhardt/chapters/5-offline_design_tezos.tex
Final revision chpt 5 until compiler

\figref{fig:interaction_monolithic} depicts the resulting interactions of the nodes, validators and monolithic smart contracts on the blockchain. The execution model requires a dedicated modus for distributed assertion checking. A possible mechanism to activate this modus is using a special transaction type, which is denoted by \texttt{txa(destination, entrypoint, parameter)} in the graphic. When an operation, such as a transaction or origination, is injected into a node, it is normally propagated to the network after it has been pre-validated by the node. After the injection of \texttt{txa(1, \%A, parameter)}, however, the node broadcasts its derivation \texttt{txa(1, \%A\_c, parameter)} to the network, which invokes the controller entrypoint instead. If this operation finishes successfully for all validators, hence no counterexample is published on the network after a certain amount of block cycles, the node can finally broadcast the initial operation as a normal transaction to the network. In case of an assertion failure, the respective validator broadcasts the counterexample with the generated value to the network. As a consequence of the broadcasted counterexample, the initial operation \texttt{txa(1, \%A, parameter)} should be reverted and rejected by the injecting node. How a counterexample is represented and included on the chain is an unresolved question --- \secref{sec:counterexample} picks this topic up and gives some ideas regarding possible implementations.

\begin{figure}
\centering
  \includegraphics[width=0.7\textwidth]{figures/5-offline_tezos/interaction_monolithic_2}
	\caption{Interaction with a monolithic contract}
	\label{fig:interaction_monolithic}
\end{figure}

This orchestration scheme only works as described when using explicit entrypoint tags for the contract invocations. On the one hand, calling entrypoints implicitly (by wrapping the parameter in the respective sum type constructors) causes a type mismatch for the controller and assertion entrypoints. \lstref{lst:mono} shows the parameter type of the Michelson implementation corresponding to the contract given in \lstref{lst:mono_liq} --- entrypoint \texttt{A} can be invoked implicitly with parameter value \texttt{Left 4}. The respective assertion entrypoint \texttt{A\_a} expects the value as \texttt{Right (Right (Left 4))} and the controller \texttt{A\_c} as \texttt{Right (Right (Right (Left 4)))}. Calling these entrypoints with the same parameter is thus not possible. \\
On the other hand, the explicit tags encode the link between production, controller and assertion code --- omitting them would dissolve all references between them. Hence, implicit entrypoint invocation cannot be supported in the execution mode of assertions. The compiler should enforce the use of explicit entrypoints in the parameter type declarations and the \texttt{txa} operation type must take a mandatory entrypoint specification as an argument.

This thesis does not go into more detail about how the described execution model for a monolithic orchestration is designed and implemented in the blockchain protocol. What needs to be considered in future work concerning the protocol design, is the separate execution mode, possibly a dedicated transaction type as proposed above, the derivation of a transaction which attempts to find a counterexample, the handling of assertion failures and the exclusive acceptance of either the parameter (and thus the transaction) or a counterexample.

\lstinputlisting[label=lst:mono, language=Michelson, caption=Skeleton of a monolithic contract in Michelson, numbers=left, float]{listings/monolithic.tz}

\subsubsection{Evaluation of the Monolithic Approach}
The presented implementations of the monolithic approach each have strengths and weaknesses. Generally, they share a pivotal disadvantage - they require a modification of the parent contract. This complicates the compilation process substantially and precludes the integration of assertions to contracts which have already been deployed to the blockchain. \lstref{lst:mono} shows how the parameter type and branching within the code body have to be adapted compared to the original parent contract in \lstref{lst:mono_parent}. Since the monolithic assembly results in deeper parameter and code nestings, the overall complexity of the contract increases and may obscure the production code.
\lstinputlisting[label=lst:mono_parent, language=Michelson, caption=Skeleton of the pure parent contract in Michelson, numbers=left]{listings/monolithic_parent.tz}

@@

Besides these disadvantages, the monolithic approach has some distinct advantages. The production and controller entrypoints can be invoked independently, implicating that the validators do not have to run the production code while checking the assertion. Conversely, the assertion code is not rerun if the parameter is valid and the transaction is broadcast to the network. This reduces the transaction cost for the user, as well as the overall computational footprint of the transaction in the network.

\subsection{Modular Orchestration}\label{sec:modular}
In the modular approach, the controller and each of the assertions are originated as separate contracts on the blockchain. The controller mirrors the interface of the parent contract, while each assertion contract only features a single entrypoint which corresponds to the raw type of the associated parent entrypoint. By keeping the assertion code separated, the parent contract remains pure and does not have to be modified. Similar to the hybrid solution mentioned in the monolithic approach, the controller calls the respective assertion contract $t$ times as an internal operation, before invoking the parent contract as operation $t+1$. In the case of empty assertions, the parent contract is invoked immediately --- the controller thus acts as a direct gateway. Corresponding to this architecture, the toolchain returns $m + 1$ contracts for a assertion specification, where $m$ is the number of specified assertions in the input. The code assembly and output of the toolchain is outlined in \figref{fig:modular_assembly}.
\begin{figure}[h]
\centering
  \includegraphics[width=0.6\textwidth]{figures/5-offline_tezos/pipeline_output_modular.png}
	\caption{Modular assembly of assertion and manager code}
	\label{fig:modular_assembly}
\end{figure}

In order to link the separate contracts to each other, the controller needs to have a reference to the other contract modules. This results in a fixed origination order --- the assertion and parent contracts must be originated first, and their addresses are passed to the controller as initial storage values. A proposed structure of the storage is a map from a unique identifier to an address. The entrypoints of the controller can then look up the address with their assigned identifier. The assignment of identifiers to entrypoints is hard-wired during the compilation, as is demonstrated in the rudimentary and simplified Liquidity implementation of a controller contract in \lstref{lst:manager_modular}. The corresponding storage initiation is the map $\lbrace 0 \rightarrow \langle \text{addr parent} \rangle; 1 \rightarrow \langle \text{addr assertion} \rangle \rbrace$ . The hard-wired identifiers are used in lines 5 and 6 to retrieve the addresses from the storage.
\lstinputlisting[label=lst:manager_modular, caption=Implementation of the modular controller in simplified Liquidity, language=Michelson, numbers=left]{listings/controller_basic.liq}

The interaction with the modular contracts is shown in \figref{fig:interaction_modular}. Even though the transactions shown in the diagram state an explicit entrypoint tag, the scheme also supports implicit entrypoint invocations. Invocations of the assertion contract pass the parameter as a raw value, i.e., unpacked from the sum type constructors. The general execution model is similar to the model for the monolithic approach, but requires a few modifications; because the controller mirrors the interface of the parent contract, the \texttt{txa} operation can be broadcast to the network directly --- a derivation to a distinct transaction, which calls the assertion selectively, is not necessary. If none of the validators publish a counterexample, the initial transaction is broadcasted to the network as a normal operation to be included in a block. This transaction will inevitably rerun the assertion code in the normal execution mode (as it invokes the controller). Hence, the \texttt{random} instruction should not cause an internal operation to fail in the normal execution mode of the VM. This could be solved by marking these internal operations with the operation result ``Skipped'' (cf. \secref{sec:internal_ops}), or by introducing a dedicated result type for this case. 
\begin{figure}[h]
\centering
  \includegraphics[width=0.9\textwidth]{figures/5-offline_tezos/interaction_modular.png}
	\caption{Interaction with modular contracts}
	\label{fig:interaction_modular}
\end{figure}

@@

The approach essentially has two weaknesses: firstly, the assertion and the parent cannot be called separately. As a result, the production code is run by every validator while searching for counterexamples, and when validating a block containing the accepted transaction. This increases the transaction cost for the user and requires a higher computational effort by the whole network. Secondly, it exposes a security vulnerability, given that assertions can be bypassed by calling the parent contract directly. This iteration of the design does not demand to close this vulnerability, hence possible solutions are not discussed in this thesis.

@@

\section{Backend Implementation}\label{sec:backend_impl}
The following sections describe the tasks and implementations of the pipeline stages of the toolchain backend.
\begin{figure}[h]
\includegraphics[width=\linewidth]{figures/5-offline_tezos/pipeline_backend}
\caption{The stages of Tezos-specific backend of the compilation pipeline}
\label{fig:pipeline_backend}
\end{figure}

\subsection{Interaction with Tezos Nodes}
Some stages of the backend have to interact with a Tezos node, e.g. to retrieve information from the blockchain. For this, the backend uses the Tezos client libraries, which provide high-level functions wrapping RPCs to a configured node. The command line interface of the toolchain provides options for configuring the host address and port of the RPC interface to connect to a specific local or public node. By default, the toolchain tries to connect to a local node on the default port.

@@

Semantic rules imposed by Michelson do not have to be checked explicitly by the toolchain --- the RPC interface of Tezos nodes provides a function to parse and type check Michelson programs, which can be used to verify the correctness of the compilation result and reject any assertions with containing typing or other semantic errors.

If the semantic check was successful, the module returns the assertions as a list of Tezos-specific ASTs. \lstref{lst:tezos_ast_type} shows the toplevel definition of this AST type.
\begin{lstlisting}[label=lst:tezos_ast_type, language=, numbers=none, caption=Toplevel Tezos-specific AST type definition in OCaml]
type tezos_ast =
  {entrypoint : string * pattern; body : assertion}
\end{lstlisting}
As Tezos is currently the only supported target platform, the type differs only slightly from the generic type. If the assertion does not specify an explicit entrypoint tag, the optional value is replaced with Tezos' notation for the default entrypoint. The subtree types for the parameter pattern and the assertion remain equivalent.

\subsection{Type Check}\label{sec:typecheck}
Given a list of Tezos-specific ASTs, the task of the type checker is to link each AST to the respective entrypoint of the parent contract. This stage module can be alternatively referred to as the linker. The linking is needed by the compiler to correctly assemble and type the controller and assertion contracts. As a simple example, the linker should correctly assign the AST of the first assertion in \texttt{Example.tza} (\lstref{lst:linker_assertion}) to entrypoint A of the parent contract \texttt{Example.tz} ( \lstref{lst:linker_parent}), and the second assertion to entrypoint B.

@@

Because explicit entrypoint tags are optional in Michelson contracts, the linking is primarily done by type matching between the declared parameter types in the parent and the assertions. Tags, if specified, are only considered if type matching does not result in an unambiguous assignment. The linking fails, if an assertion cannot be matched to any parent entrypoint, or several assertions match the same entrypoint, i.e., if no injective, non-surjective mapping was found.

As a first step, the linker retrieves and parses the parameter type of the parent contract. The parent contract is provided by the user as a file, a blockchain address, or a raw string. In case the contract is given as an address, the code is retrieved from the blockchain. The parameter pattern of each assertion AST is then compared recursively to the types of each entrypoint, until a match has been found --- this entrypoint is then blocked for other assertions.

As explained in \secref{sec:tezos}, entrypoints can be invoked implicitly by wrapping the parameter into the respective union type constructors. Alternatively, they can be called explicitly by stating an entrypoint tag and passing the raw parameter value. The same applies for the specification of assertions - when omitting the tag, the parameter pattern must be declared in respect to the default entrypoint. If the assertion states an explicit tag, it may declare a pattern of the raw parameter type. \lstref{lst:impl_expl} shows two valid assertion specifications for entrypoint A of \lstref{lst:linker_parent} --- the first uses an explicit and the second an implicit declaration of the parameter pattern. 
\begin{lstlisting}[language=Assertion, numbers=none, caption=Explicit and implicit assertion specification, label=lst:impl_expl]
(entrypoint %A (i : int) ...)
(entrypoint (left (i : int)) ...)
\end{lstlisting}

The tags given in the assertion specification do not necessarily have to be identical with the tags of the parent contract; in some cases, they may be chosen differently for documentation purposes. However, the tags can and should be used to resolve ambiguity in the mapping between assertions and entrypoints. \lstref{lst:ambiguity_assertion} provides four examples of assertions, which cannot be assigned unambiguously to an entrypoint of the contract given in \lstref{lst:ambiguity_parent}. Since the entrypoints of \texttt{Example2.tz} have the same raw parameter type, and none of the given assertions tags refer to one of them explicitly, the linker cannot infer which entrypoint they refer to. In all four cases, providing the tag \texttt{A} or \texttt{B} resolves the ambiguity. If the parent entrypoint does not specify a tag, the ambiguity can be resolved by specifying the parameter type in respect to the default entrypoint, which is unambiguous by design.

@@

Another type of ambiguity can be caused by the fact that entrypoints can be sub-entrypoints of others. As an example, consider the parameter type \texttt{(or (int \%A) (or \%BC (int \%B) (int \%C)))}. Entrypoints \texttt{B} and \texttt{C} are sub-entrypoints of \texttt{BC} --- invoking \texttt{BC} will inevitably result in an invocation of either \texttt{B} or \texttt{C}, depending on the parameter value. Thus, assertions may ``overlap'' if a separate assertion is declared for a super- and a sub-entrypoint and result in two assertion specifications for the same entrypoint. Hence, the linker needs to detect overlapping assertions and reject assertion contracts where appropriate.

In summary, the linker rejects any input containing ambiguous, duplicate, overlapping or mismatching assertions, i.e., assertions whose types do not match any entrypoint. If each assertion could be linked to exactly one entrypoint, the mapping from entrypoint to assertion is returned. Because entrypoints can be anonymous, the mapping represents the entrypoints by the constructor nesting of their parameters. The resulting mapping for \texttt{Example.tz} and \texttt{Example.tza}, for instance, would be \\
\texttt{\{ Left \textrightarrow <AST assertion A>; Right \textrightarrow <AST assertion B>\} }.

@@

With Michelson as the compilation target language, the compiler has to generate code for a stack machine. After compilation, the Michelson programs can be validated for correctness by the Tezos node, which will intercept any assertions containing type errors. After sanity-checking the programs, the output of the compiler is the target code for the controller and assertion contracts. As an additional output, it should return a piece of template code for the correct storage initialization of the controller. Before originating the controller contract, the user fills in the blockchain addresses of the other modules.

@@

If the pattern excludes any special values of the data type (e.g., \texttt{cons head tail} excludes empty lists), the generated code should reflect this. Excluded values should never cause the assertion to fail. In the example of \texttt{cons}, the generated code should thus contain the branching shown in \lstref{lst:cons}. The branch in line 1 is executed if the list contains at least one element, and should contain the assertion code. If the list is empty, the branch in line 2 should simply return successfully. 
\begin{lstlisting}[language=Michelson, label=lst:cons, caption=Exclusion of empty lists in the assertion]
IF_CONS ( <assertion body> )
        ( <return ([], ())>)
\end{lstlisting}

The generated code of the assertion body and the parameter type can be plugged into a template, which contains the necessary boilerplate code for a contract. Such a template is shown in \lstref{lst:contract_templ} --- the placeholders \texttt{<?>} mark the locations where the individual code is inserted. Lines 3-5 prepare the stack, s.t. the parameter and storage are the two top elements. The lines below the placeholder for the body clean up the stack and leave it according to calling convention, i.e., the program returns an empty list of internal operations and the unit storage.
\lstinputlisting[label=lst:contract_templ, language=Michelson, caption=Template code for assertion contracts in Michelson, numbers=left]{listings/assertion_template.tz}

\subsubsection{Compilation of the Controller Contract}
The generation of the \texttt{parameter} and \texttt{storage} primitives is trivial. Given that the controller acts as a proxy, the parameter type can be copied from the parent contract. Assuming that the proposal for the storage type from \lstref{lst:manager_modular} is adopted, the storage type is always the same, i.e., \texttt{(map int address)}. The code body is generated recursively based on the parameter type. It consists of two types of entrypoints, which either act as a direct gateway to the parent contract, or trigger an assertion check before invoking the parent.

\paragraph{Gateway entrypoints}
The code of gateway entrypoints only has to execute two steps; first, the address of the parent contract is read from the storage. It then creates an internal operation to this address, passing along the original parameter and funds. Assuming the parent address is always stored under the same key, e.g. with id 0, such gateways can be inserted as fixed boilerplate code.

\paragraph{Assertion entrypoints}
If the entrypoint should trigger an assertion, the compiler has to generate code executing the following steps:
\begin{enumerate}
\itemsep-0.5em
\item Read the addresses from the assertion and parent contract from storage
\item Calculate $t$
\item Compile a list of $t$ internal operations invoking the assertion with the raw parameter
\item Create an internal operation calling the parent contract, passing the original parameter and funds
\item Return a list of all operations and the unchanged storage
\end{enumerate}
\secref{sec:prob_threshold} stated a formula to determine the total number $t_{total}$ of required test runs for a given probability threshold $c$, and the size of the search space $n$. From $t_{total}$, the number of test runs $t$ per validator can be calculated according to \eqref{eq:t_validator}. Since $n$ can only be obtained at runtime, the compiler needs to derive a formula to determine this value from the ranges of the random generators. \secref{sec:sample_space} already gave some examples of this. 
The threshold $c$, on the other hand, must be given as a parameter to the compiler. Therefore, the command line interface of the toolchain should provide an option to set this threshold.

>>>>

\figref{fig:interaction_monolithic} depicts the resulting interactions of the nodes, validators and monolithic smart contracts on the blockchain. The execution model requires a dedicated mode for distributed assertion checking. A possible mechanism to activate this mode is using a special transaction type, which is denoted by \texttt{txa(destination, entrypoint, parameter)} in the graphic. When an operation, such as a transaction or origination, is injected into a node, it is normally propagated to the network after it has been pre-validated by the node. After the injection of \texttt{txa(1, \%A, parameter)}, however, the node broadcasts its derivation \texttt{txa(1, \%A\_c, parameter)} to the network, which invokes the controller entrypoint instead. If this operation finishes successfully for all validators, i.e., no counterexample is published on the network after a certain amount of block cycles, the node can finally broadcast the initial operation as a normal transaction. In case of an assertion failure, the respective validator broadcasts the counterexample including the generated random values. As a consequence of the broadcasted counterexample, the initial operation \texttt{txa(1, \%A, parameter)} should be reverted and rejected by the injecting node. How a counterexample is represented and included on the chain is an unresolved question --- \secref{sec:counterexample} picks this topic up and gives some ideas regarding possible implementations.

\begin{figure}
\centering
  \includegraphics[width=0.9\textwidth]{figures/5-offline_tezos/interaction_monolithic_2}
	\caption{Interaction with a monolithic contract}
	\label{fig:interaction_monolithic}
\end{figure}

This orchestration scheme only works as described when using explicit entrypoint tags for the contract invocations. On the one hand, calling entrypoints implicitly (by wrapping the parameter in the respective sum type constructors) causes a type mismatch for the controller and assertion entrypoints. \lstref{lst:mono} shows the parameter type of the Michelson implementation corresponding to the contract given in \lstref{lst:mono_liq} --- entrypoint \texttt{A} can be invoked implicitly with parameter value \texttt{Left 4}. The respective assertion entrypoint \texttt{A\_a} expects the value as \texttt{Right (Right (Left 4))} and the controller \texttt{A\_c} as \texttt{Right (Right (Right (Left 4)))}. Calling these entrypoints with the same parameter is thus not possible. On the other hand, the explicit tags encode the link between production, controller and assertion code --- omitting them would dissolve all references between them. Hence, implicit entrypoint invocation cannot be supported in the execution mode of assertions. The assertion compiler thus needs to enforce the use of explicit entrypoints in the parameter type declarations and the \texttt{txa} operation type must expect a mandatory entrypoint specification as an argument.
\lstinputlisting[label=lst:mono, language=Michelson, caption=Skeleton of a monolithic contract in Michelson, numbers=left, float]{listings/monolithic.tz}

This thesis does not go into more detail about how the described execution model for a monolithic orchestration is designed and implemented in the blockchain protocol. What needs to be considered in future work concerning the protocol design is the separate execution mode, a dedicated operation type as proposed above, the derivation of a transaction which attempts to find a counterexample, the handling of assertion failures and the exclusive acceptance of the off-chain computation (and thus the transaction submitting it) or the counterexample.

\subsubsection{Evaluation of the Monolithic Approach}
The purely and hybrid monolithic implementation each have strengths and weaknesses. Generally, they share a pivotal disadvantage - they require a modification of the parent contract. This complicates the compilation process substantially and precludes the integration of assertions to contracts which have already been deployed to the blockchain. \lstref{lst:mono} shows how the parameter type and branching within the code body have to be adapted compared to the original parent contract in \lstref{lst:mono_parent}. Since the monolithic assembly results in deeper parameter and code nestings, the overall complexity of the contract increases and may obscure the production code.
\lstinputlisting[label=lst:mono_parent, language=Michelson, caption=Skeleton of the pure parent contract in Michelson, numbers=left]{listings/monolithic_parent.tz}

@@

Besides these disadvantages, the monolithic approach provides a distinct advantage: the production and controller entrypoints can be invoked independently, implicating that the validators do not have to run the production code while checking the assertion. Conversely, the assertion code is not rerun if the parameter was validated and the actual transaction is broadcast to the network. This reduces the transaction cost for the user, as well as the overall computational footprint of the transaction and the validation in the network.

\subsection{Modular Orchestration}\label{sec:modular}
In the modular approach, the controller and each of the assertions are originated as separate contracts on the blockchain. The controller mirrors the interface of the parent contract, while each assertion contract only features a single entrypoint, which corresponds to the raw type of the associated parent entrypoint. By keeping the assertion code separated, the parent contract remains pure and does not have to be modified. Similar to the hybrid solution mentioned in the monolithic approach, the controller calls the respective assertion contract $t$ times as an internal operation, before invoking the parent contract as operation $t+1$. In the case of empty assertions, the controller acts as a direct gateway by invoking the parent contract immediately. Corresponding to this architecture, the toolchain returns $m + 1$ contracts, where $m$ is the number of specified assertions in the input. The code assembly and output of the toolchain is outlined in \figref{fig:modular_assembly}.
\begin{figure}[h]
\centering
  \includegraphics[width=0.6\textwidth]{figures/5-offline_tezos/pipeline_output_modular.png}
	\caption{Modular assembly of assertion and manager code}
	\label{fig:modular_assembly}
\end{figure}

In order to link the separate contracts to each other, the controller needs to have a reference to the other contract modules. This results in a fixed origination order --- the assertion and parent contracts must be originated first, and their addresses are passed to the controller as initial storage values. A proposed structure of the storage is a map from a unique identifier to an address. The entrypoints of the controller can then look up the addresses of the parent and the respective assertion contract with an assigned identifier. The assignment of the identifiers is hard-wired during the compilation, as is demonstrated in the rudimentary and simplified Liquidity implementation of a controller contract in \lstref{lst:manager_modular}. The corresponding storage initiation for this contract is the map $\lbrace 0 \rightarrow \langle \text{address parent} \rangle; 1 \rightarrow \langle \text{address assertion} \rangle \rbrace$ . The hard-wired identifiers are used in lines 5 and 6 to retrieve the addresses from the storage.
\lstinputlisting[label=lst:manager_modular, caption=Implementation of the modular controller in simplified Liquidity, language=Michelson, numbers=left]{listings/controller_basic.liq}

The interaction with the modular contracts is shown in \figref{fig:interaction_modular}. Even though the transactions shown in the diagram state an explicit entrypoint tag, the scheme also supports implicit entrypoint invocations. Invocations of the assertion contract pass the parameter as a raw value. The general execution model is similar to the model for the monolithic approach, but requires a few modifications; because the controller mirrors the interface of the parent contract, the \texttt{txa} operation can be broadcast to the network directly --- a derivation to a distinct transaction, which calls the assertion selectively, is not necessary. If none of the validators publish a counterexample, the initial transaction is broadcasted to the network as a normal operation to be included in a block. This transaction will inevitably rerun the assertion code in the normal execution mode (as it invokes the controller). Hence, the \texttt{random} instruction should not cause an internal operation to fail in the normal execution mode of the VM. This could be solved by marking these internal operations with the operation result ``Skipped'' (cf. \secref{sec:internal_ops}), or by introducing a dedicated result type for this case. 
\begin{figure}[h]
\centering
  \includegraphics[width=\textwidth]{figures/5-offline_tezos/interaction_modular.png}
	\caption{Interaction with modular contracts}
	\label{fig:interaction_modular}
\end{figure}

@@

The approach essentially exhibits two weaknesses: firstly, the assertion and the parent cannot be called separately. As a result, the production code is run by every validator while searching for counterexamples. Moreover, the assertion code is also run when validating a block containing the accepted transaction. This increases the transaction cost for the user and requires a higher computational effort by the whole network. Secondly, it exposes a security vulnerability, given that assertions can be bypassed by calling the parent contract directly. This iteration of the design does not demand to close this vulnerability, hence possible solutions are not discussed in this thesis.

@@

\section{Backend Implementation}\label{sec:backend_impl}
The following sections describe the tasks and implementation of the pipeline stages in the toolchain backend.
\begin{figure}[h]
\includegraphics[width=\linewidth]{figures/5-offline_tezos/pipeline_backend}
\caption{The stages of Tezos-specific backend of the compilation pipeline}
\label{fig:pipeline_backend}
\end{figure}

\subsection{Interaction with Tezos Nodes}
Some stages of the backend have to interact with a Tezos node, e.g. to retrieve information from the blockchain. For this, the backend uses the Tezos client libraries, which provide high-level functions wrapping RPCs to a configured node. The command-line interface of the toolchain provides options for configuring the host address and port of the RPC interface to connect to a specific local or public node. By default, the toolchain tries to connect to a local node on the default port.

@@

Semantic rules imposed by Michelson do not have to be checked explicitly by the toolchain --- the RPC interface of Tezos nodes provides a function to parse and type check Michelson programs, which can be used to verify the correctness of the compilation result and reject any assertions containing type or other semantic errors.

If the semantic check was successful, the module returns the assertions as a list of Tezos-specific ASTs. \lstref{lst:tezos_ast_type} shows the toplevel definition of this AST type.
\begin{lstlisting}[label=lst:tezos_ast_type, language=, numbers=none, caption=Toplevel Tezos-specific AST type definition in OCaml]
type tezos_ast =
  {entrypoint : string * pattern; body : assertion}
\end{lstlisting}
As Tezos is currently the only supported target platform, the type differs only slightly from the generic type. If the assertion does not specify an explicit entrypoint tag, the optional value is replaced with Tezos' notation for the default entrypoint. The subtree types for the parameter pattern and the assertion remain equivalent to the generic subtree types.

\subsection{Type Check}\label{sec:typecheck}
Given a list of Tezos-specific ASTs, the task of the type checker is to link each AST to the respective entrypoint of the parent contract. This stage module can therefore also be referred to as the ''linker``. The linking is needed by the compiler to correctly assemble and type the controller and assertion contracts. As a simple example, the linker should correctly assign the AST of the first assertion in \texttt{Example.tza} (\lstref{lst:linker_assertion}) to entrypoint A and the second assertion to entrypoint B of the parent contract \texttt{Example.tz} (\lstref{lst:linker_parent}).

@@

Because explicit entrypoint tags are optional in Michelson contracts, the linking is primarily done by type matching between the declared parameter types in the parent and parameter patterns in the assertions. If the type matching does not result in an unambiguous assignment between an assertion and an entrypoint, i.e., the parameter pattern matches the type of several entrypoints, the linker considers the specified tags in an effort to resolve the ambiguity. The linking fails if no injective, non-surjective mapping from the set of assertions to the set of entrypoints was found regardless.

\subsubsection{Parameter Patterns Revisited}
As explained in \secref{sec:tezos}, transactions on Tezos can invoke entrypoints implicitly by invoking the default entrypoint and wrapping the parameter value into the respective union type constructors. Alternatively, they can be called explicitly by stating an entrypoint tag and passing the raw parameter value as an argument. The same principle applies for the specification of assertions - when omitting the tag in the assertion signature, the parameter pattern must be declared in respect to the default entrypoint. If the assertion states an explicit tag, it may declare a pattern of the raw parameter type. \lstref{lst:impl_expl} shows two valid assertion specifications for entrypoint A of the contract \texttt{Example.tz} (\lstref{lst:linker_parent}) --- the first specification uses an explicit tag paired with the pattern of the raw parameter type, while the second does not specify a tag and declares the parameter pattern with respect to the default entrypoint, i.e, wrapped into the sum type constructor \texttt{left}. 
\begin{lstlisting}[language=Assertion, numbers=none, caption=Explicit and implicit parameter specification in assertions, label=lst:impl_expl]
(entrypoint %A (i : int) ...)
(entrypoint (left (i : int)) ...)
\end{lstlisting}

The tags given in the assertion specification do not necessarily have to be identical with the tags of the parent contract; they may be chosen differently for documentation purposes, particularly if the parent contract does not state any tags. However, tags can and should be used to resolve ambiguity in the linking between assertions and entrypoints by mirroring the entrypoint tags of the parent contract.

\subsubsection{Resolving ambiguity}
\lstref{lst:ambiguity_assertion} provides four examples of assertions which cannot be linked unambiguously to an entrypoint of the contract given in \lstref{lst:ambiguity_parent}. Since the entrypoints of \texttt{Example2.tz} have the same raw parameter type \texttt{int}, and none of the given assertions tags refer to one of them explicitly, the linker cannot infer which entrypoint they refer to. In all four cases, providing an explicit tag (\texttt{A} or \texttt{B}) can resolve the ambiguity. If the parent entrypoint does not specify a tag, the ambiguity can be resolved by omitting the tag and specifying the parameter pattern in respect to the default entrypoint, e.g. \texttt{left \_} or \texttt{right (i : int)}.

@@

Another type of ambiguity is caused by the fact that entrypoints can be sub-entrypoints of others. As an example, consider the parameter type of contract \texttt{Example3} in \lstref{lst:overlap_parent}. Entrypoints \texttt{B} and \texttt{C} are sub-entrypoints of \texttt{BC} --- invoking \texttt{BC} will inevitably result in an invocation of either \texttt{B} or \texttt{C}, depending on the parameter value. Thus, assertions may ``overlap'' if a separate assertion is declared for a super- and a sub-entrypoint and result in two assertion specifications for the same entrypoint, as is demonstrated in \lstref{lst:overlap_assertion}. The linker detects such overlapping assertions and rejects assertion specifications containing such. Hence, a valid assertion contract can specify an assertion for entrypoint \texttt{AB} only, or for entrypoints \texttt{A} and \texttt{B}, but not for \texttt{AB}.

\vspace{\baselineskip}
\noindent
\begin{minipage}{.45\textwidth}
\begin{lstlisting}[label=lst:overlap_parent, numbers=none, language=Michelson, caption=Example3.tz]
parameter (or %AB (int %A)
                  (int %B));
...
\end{lstlisting}
\end{minipage}\hfill
\begin{minipage}{.5\textwidth}
\begin{lstlisting}[label=lst:overlap_assertion, numbers=none, language=Assertion, caption=Example3.tza with overlapping assertions]
(entrypoint %AB (left  i) ...)
(entrypoint %A j ...)
\end{lstlisting}
\end{minipage}
\vspace{\baselineskip}

\subsubsection{Linking Process}
As a first step, the linker retrieves and parses the parameter type of the parent contract. The parent contract is provided by the user as a file, a blockchain address, or a raw string. In case the contract is given as an address, the code is retrieved from the blockchain. From the parsed parameter type, a list of all named and anonymous entrypoints with their raw input types is extracted. For these steps, the Tezos client and protocol libraries are utilized.

Based on the extracted pool of entrypoints and the list of assertion ASTs, the linker then iteratively builds a mapping between entrypoints and assertions. The parameter pattern of each assertion AST is compared to the parameter type of all entrypoints in the pool. All matching entrypoints are collected in a list. If this list is empty, linking the assertion fails due to a type mismatch and the pipeline terminates. In case the list contains several entrypoints, the linker considers the tags in order to resolve the ambiguity. If the tags do not lead to a resolution, the pipeline terminates due to an ambiguous assertion specification. Lists containing exactly one match can be resolved to an assignment directly. After an AST could be linked to an entrypoint, the assignment is added to the mapping and the matching entrypoint is removed from the pool of unlinked entrypoints. Additionally, all sub- and super-entrypoints are removed from the pool. By removing these entrypoints, the linker is able to detect duplicate or overlapping assertions.

After successfully finding a valid assignment for all assertions, the linker returns the mapping. Because entrypoints can be anonymous, the mapping represents the entrypoints with the constructor nesting of their parameters. \lstref{lst:unionpath} shows the definition of the type for such a representation.
\begin{lstlisting}[label=lst:unionpath, language=, numbers=none, caption=Type representing entrypoints according to the union constructor nesting]
type union_path = Left of union_path
                | Right of union_path
                | T
\end{lstlisting}
Using this type, entrypoints \texttt{A} and \texttt{B} of \texttt{Example.tz}, for instance, are represented by \texttt{Left T} and \texttt{Right T}, while the default entrypoint is represented by \texttt{T}. Given \texttt{Example.tza}, the linker thus returns the mapping \\
\texttt{\{ Left T \textrightarrow \, <AST assertion A>; Right T \textrightarrow \, <AST assertion B>\} }.

@@

With Michelson as the compilation target language, the compiler has to generate code for a stack machine. After compilation, the Michelson programs can be validated for correctness by the Tezos node, which will intercept any assertions containing type errors. After this sanity-check, the compiler outputs the target code for the controller and assertion contracts. As an additional output, it should return a piece of template code for the correct storage initialization of the controller contract. Before originating the controller contract, the user fills in the blockchain addresses of the other modules.

@@

If the pattern excludes any special values of the data type (e.g., \texttt{cons head tail} excludes empty lists), the generated code should reflect this. It has not been specified yet, if excluded values should never or always cause the assertion to fail. In the example of \texttt{cons} and assuming excluded values pass the assertion, the generated code could contain the branching shown in \lstref{lst:cons}. The branch in line 1 is executed if the list contains at least one element, and should contain the assertion code. If the list is empty, the branch in line 2 should simply return successfully. 
\begin{lstlisting}[language=Michelson, label=lst:cons, caption=Exclusion of empty lists in the assertion]
IF_CONS ( <assertion body> )
        ( <return ([], ())>)
\end{lstlisting}

To simplify the compilation, the generated code of the assertion body and the parameter type can be plugged into a template containing the necessary boilerplate code for a contract. Such a template is shown in \lstref{lst:contract_templ} --- the placeholders \texttt{<?>} mark the locations where the individual code is inserted. Since assertions do not use any storage as of now, the storage type is set to unit. Lines 3-5 prepare the stack, s.t. the parameter and storage are the two top elements. The lines below the placeholder for the body leave the stack according to calling convention, i.e., the program returns an empty list of internal operations and the unit storage.
\lstinputlisting[label=lst:contract_templ, language=Michelson, caption=Template code for assertion contracts in Michelson, numbers=left]{listings/assertion_template.tz}

\subsubsection{Compilation of the Controller Contract}
The generation of the \texttt{parameter} and \texttt{storage} primitives for the controller contract is trivial. Given that the controller acts as a proxy, the parameter type can be copied from the parent contract. Assuming that the proposal for the storage type from \lstref{lst:manager_modular} is adopted, the storage type is always the same, i.e., \texttt{(map int address)}. The code body is generated recursively based on the parameter type. It consists of two types of entrypoints, which either act as a direct gateway to the parent contract in the case of empty assertions, or trigger an assertion check before invoking the parent.

\paragraph{Gateway entrypoints}
The code of gateway entrypoints only has to execute two steps; first, the address of the parent contract is read from the storage. It then creates an internal operation to this address, passing along the original parameter and funds. Assuming the parent address is always stored under the same key, e.g. with id 0, such gateways can be generated as fixed boilerplate code.

\paragraph{Assertion entrypoints}
If the entrypoint should trigger an assertion, the compiler has to generate code executing the following steps:
\begin{enumerate}
\itemsep-0.5em
\item Read the addresses from the assertion and parent contract from storage
\item Calculate the number of test runs $t$
\item Compile a list of $t$ internal operations invoking the assertion with the raw parameter
\item Create an internal operation calling the parent contract, passing the original parameter and funds
\item Return (i.e., push to the stack) a list of all internal operations and the unchanged storage
\end{enumerate}
\secref{sec:prob_threshold} stated a formula to determine the total number $t_{total}$ of required test runs, given a probability threshold $c$ and the size of the search space $n$. From $t_{total}$ the number of test runs $t$ per validator can be calculated according to \eqref{eq:t_validator}. Since $n$ can only be obtained at runtime, the compiler needs to derive a formula to determine this value from the ranges of the random generators. \secref{sec:sample_space} already gave some examples of this. 
The threshold $c$, on the other hand, must be given as a parameter to the compiler. Therefore, the command line interface of the toolchain should provide an option to set this threshold.
